---
title: "Homework - 1"
author: "24203476 - Vyom Pandey"
format: pdf
editor: visual
execute: 
  echo: false
---

# Introduction

The `airquality` dataset is a *built-in* R dataset that contains 'daily measurements of air quality in New York' from 'May to September 1973'. In this analysis, we examine how **temperature** influences **ozone levels**—a critical factor for public health and environmental monitoring.

This relationship is especially relevant in the context of **climate change**, where rising global temperatures could lead to more frequent air quality alerts and health advisories.

# Descriptive Statistics

We compute `summary statistics` for the two variables of interest: **Ozone** and **Temperature**.

```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
```

# Clean dataset

```{r, warning=FALSE, message=FALSE}
data <- na.omit(airquality)
# Remove all rows with missing values (NA) from the airquality dataset
summary_stats <- data %>%
  summarise(
    Count = n(),    # Total number of complete observations (rows)
    Mean_Ozone = mean(Ozone), # Mean (average) of Ozone levels
    SD_Ozone = sd(Ozone), # Standard deviation of Ozone levels
    Mean_Temp = mean(Temp), # Mean (average) of Temperature
    SD_Temp = sd(Temp)   # Standard deviation of Temperature
  )
knitr::kable(summary_stats, caption = "Table 1: Summary Statistics for Ozone and Temperature")

```

# Plot: Ozone vs Temperature

```{r, warning=FALSE}
ggplot(data, aes(x = Temp, y = Ozone)) +
  # Create a ggplot object using the dataset `data`, mapping Temperature to the x-axis and Ozone to the y-axis.

  geom_point(color = "#2c7fb8", alpha = 0.7) +
  # Add a scatter plot (points) with a specific blue color and transparency for better visibility.

  geom_smooth(method = "lm", se = FALSE, color = "darkred") +
  # Add a linear regression line (method = "lm") without the confidence interval shading (se = FALSE) and set colour to darkred

  labs(
    title = "Ozone Concentration vs Temperature in New York (1973)",
    x = "Temperature (°F)",
    y = "Ozone (ppb)"
  ) +
  # Add descriptive labels: title of the plot, and axis labels for x and y.

  theme_minimal()
  # Apply a minimalistic theme to the plot, which reduces visual clutter for a cleaner look.

```

# Interpretation

The plot reveals a **positive correlation** between temperature and ozone levels. As temperature increases, ozone concentration also tends to rise. This trend suggests that hotter days are associated with poorer air quality, likely due to photochemical reactions that increase ozone formation. This has significant implications for public health, especially during summer months.

# Conclusion

In this report, we:

-   Analyzed the relationship between temperature and ozone levels in New York City

-   Found a clear positive trend between the two variables

-   Highlighted how environmental conditions can influence air pollution levels

The `airquality` dataset provides valuable historical insight into the impact of temperature on urban air quality.
